The Talking Heads Experiment

### 2015-04-27 chap3/chap3.tex
first pass of bibliography

Basic behaviours can be put together to form 
more complex behaviours.\footnote{A representative sample of experiments in this 
direction is reported in: \cite{Steels:1996}. See 
also: \cite{Arkin:1998}.}
\figref{phototaxis} 
shows a recording of the internal states of the robot
in \figref{f:plate3} as it is performing 
phototaxis towards a light source, using two simple
light sensors, and touch-based obstacle avoidance
behaviour.  When the robot strikes the box housing 
the light, it moves back in a kind of reflex behaviour
triggered by the touch sensors mounted at the front. 
The LeftLight and 
RightLight sensory channels and the LeftMotorSpeed
and RightMotorSpeed are shown 
in \figref{phototaxis}. When the robot strikes
the obstacle housing the light, its left and right
motor is pulled down to a negative value so that 
it moves backwards. Then the robot zig-zags to the 
light source until it hits the obstacle again.

>>>>

Basic behaviours can be put together to form 
more complex behaviours.\footnote{A representative sample of experiments in this 
direction is reported in: \cite{Steels:1995}. See 
also: \cite{Arkin:1998}.}
\figref{phototaxis} 
shows a recording of the internal states of the robot
in \figref{f:plate3} as it is performing 
phototaxis towards a light source, using two simple
light sensors, and touch-based obstacle avoidance
behaviour.  When the robot strikes the box housing 
the light, it moves back in a kind of reflex behaviour
triggered by the touch sensors mounted at the front. 
The LeftLight and 
RightLight sensory channels and the LeftMotorSpeed
and RightMotorSpeed are shown 
in \figref{phototaxis}. When the robot strikes
the obstacle housing the light, its left and right
motor is pulled down to a negative value so that 
it moves backwards. Then the robot zig-zags to the 
light source until it hits the obstacle again.

### 2015-04-27 chap4/chap4.tex
fixed remaining widows and orphans
first pass of bibliography

Selectionism is a general means of explaining the origins of 
complexity. It requires: (1) a process that 
can generate a repertoire of possible solutions in a 
basically random fashion, (2) a process for preserving
solutions so that there can be a gradual build up of 
more complex solutions, and (3) a 
selectionist force, which uses
feedback from the environment and influences preservation
so that adequate solutions are retained 
and others discarded.\footnote{
There are many excellent introductory 
accounts of selectionism. One of the best
known is: \cite{Dawkins:1995}. The 
recent application of selectionism to the automatic
derivation of computer programs clearly demonstrates
how general the principle is. See \cite{Goldberg:1989} and \cite{Koza:1997}. 
The principle has now even been applied to 
the development of computational hardware. See \cite{Sipper:1998}.}

@@

\is{Discrimination Game}Here is a game, which I call the Discrimination Game, which is useful 
to study categorisation in 
a systematic way.\footnote{
The Discrimination Game model together with the 
discrimination trees and its growth dynamics 
was presented for the first time in \cite{Steels:1996}. 
Based on this paper a new implementation of
single category discrimination was implemented by Angus McIntyre 
within the \textsc{babel} environment. Later on, Joris Van Looveren
re-implemented the use of conjunctive combinations.}
The game is played by a single agent, randomly 
drawn from a population of agents, and 
is equivalent to the conceptualisation phase of 
the guessing game. The agent perceives the scene and 
chooses a topic from the possible objects segmented in the
scene. He then uses his discrimination trees, as developed so
far, to come up with a category or a conjunction
of categories that is valid for the topic, but not for any
other object in the context. The Discrimination Game succeeds if 
the agent has found distinctive categories, otherwise the game
fails. When the game succeeds, the success counter
of the categorisers involved go up.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.40\textwidth]{chap4/figs/scene8.pdf}}
\caption{\label{scene2}Top: The scene used in Game 8. 
Shape 0 is the topic. Bottom: The discrimination trees available
for this game.}
\end{figure}
The game is reported by the commentator as follows:
\begin{verbatim}
Game 8
 a1 segments the context in 3 objects: 
    square-0, circle-1, square-2
 a1 chooses square-0 as the topic
 The discrimination game fails
\end{verbatim}
The game fails because for the two sensory channels 
for which there are discrimination trees, the values 
of the segments are all within the lower range and so 
no distinctive category or category set could be found. 
This failure stimulates
the discrimination network to expand, but any node, 
including a top node of some of the other sensory channels 
can be chosen for further expansion.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.40\textwidth]{chap4/figs/scene22.pdf}}
\caption{\label{scene3}Top: The scene used in Game 22. 
Bottom: The discrimination trees available to the
agent.}
\end{figure}
\enlargethispage{1\baselineskip}
The next example shows a game (Game 22) based on the scene 
in \figref{scene3}. The topic is the triangle, shape 0. 
The discrimination trees for 
\textsc{height} and \textsc{width} have already more than one level and 
a discrimination tree for \textsc{hpos} has developed.
The relevant sensor-scaled data for these three sensory
channels is shown in \tabref{tab:t-game22}.

>>>>

Selectionism is a general means of explaining the origins of 
complexity. It requires: (1) a process that 
can generate a repertoire of possible solutions in a 
basically random fashion, (2) a process for preserving
solutions so that there can be a gradual build up of 
more complex solutions, and (3) a 
selectionist force, which uses
feedback from the environment and influences preservation
so that adequate solutions are retained 
and others discarded.\footnote{
There are many excellent introductory 
accounts of selectionism. One of the best
known is: \cite{Dawkins:1976}. The 
recent application of selectionism to the automatic
derivation of computer programs clearly demonstrates
how general the principle is. See \cite{Goldberg:1989} and \cite{Koza:1997}. 
The principle has now even been applied to 
the development of computational hardware. See \cite{Sipper:1998}.}

@@

\is{Discrimination Game}Here is a game, which I call the Discrimination Game, which is useful 
to study categorisation in 
a systematic way.\footnote{
The Discrimination Game model together with the 
discrimination trees and its growth dynamics 
was presented for the first time in \cite{Steels:1995}. 
Based on this paper a new implementation of
single category discrimination was implemented by Angus McIntyre 
within the \textsc{babel} environment. Later on, Joris Van Looveren
re-implemented the use of conjunctive combinations.}
The game is played by a single agent, randomly 
drawn from a population of agents, and 
is equivalent to the conceptualisation phase of 
the guessing game. The agent perceives the scene and 
chooses a topic from the possible objects segmented in the
scene. He then uses his discrimination trees, as developed so
far, to come up with a category or a conjunction
of categories that is valid for the topic, but not for any
other object in the context. The Discrimination Game succeeds if 
the agent has found distinctive categories, otherwise the game
fails. When the game succeeds, the success counter
of the categorisers involved go up.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.40\textwidth]{chap4/figs/scene8.pdf}}
\caption{\label{scene2}Top: The scene used in Game 8. 
Shape 0 is the topic. Bottom: The discrimination trees available
for this game.}
\end{figure}
The game is reported by the commentator as follows:
\newpage

\begin{verbatim}
Game 8
 a1 segments the context in 3 objects: 
    square-0, circle-1, square-2
 a1 chooses square-0 as the topic
 The discrimination game fails
\end{verbatim}
The game fails because for the two sensory channels 
for which there are discrimination trees, the values 
of the segments are all within the lower range and so 
no distinctive category or category set could be found. 
This failure stimulates
the discrimination network to expand, but any node, 
including a top node of some of the other sensory channels 
can be chosen for further expansion.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.40\textwidth]{chap4/figs/scene22.pdf}}
\caption{\label{scene3}Top: The scene used in Game 22. 
Bottom: The discrimination trees available to the
agent.}
\end{figure}

The next example shows a game (Game 22) based on the scene 
in \figref{scene3}. The topic is the triangle, shape 0. 
The discrimination trees for 
\textsc{height} and \textsc{width} have already more than one level and 
a discrimination tree for \textsc{hpos} has developed.
The relevant sensor-scaled data for these three sensory
channels is shown in \tabref{tab:t-game22}.

### 2015-04-27 extra/papers-language-steels.tex
fixed index entries and extra papers, added proofreaders and typesetters
first pass of bibliography

\chapter{Appendix}
\label{c:references}

\section{Bibliography}

@@

1. \citet{Steels:94a} The `artificial life' route to `artificial intelligence'. Building Situated Embodied Agents. Lawrence Erlbaum Ass, New Haven, 1994.

@@

9. \citet{Steels:97c} The Origins of Syntax in Visually Grounded Robotic Agents. In Pollack, M. (ed.) Proceedings of the 15th International Joint Conference on Artificial Intelligence, Morgan Kaufmann Publishers, San Francisco, CA. pp. 1632-1641.

@@

13. \citet{Steels:97g} Ancrage de Jeux de Langage Adaptatifs dans des Agents Robotiques. In A. Cherif. J. and Signorini (ed.) Intelligence Artificielle et Complexite, Paris, 1997. pp. 27-39.

@@

17. \citet{Steels:98c} The origins of syntax in visually grounded robotic agents. Artificial Intelligence, 103:1-24.

@@

21. \citet{mcintyre:99a} Net-mobile embodied agents. Proceedings of Sony Research Forum, Tokyo, 1999.

@@

28. \citet{Steels:99h} Collective learning and semiotic dynamics. In Floreano, D. and 
Nicoud, J-D and Mondada, F., editor, Advances in Artificial Life (ECAL 99) LNCS 1674, Springer Verlag, Berlin. 
pp. 679-688.

@@

31. \citet{Steels:00c} The Emergence of Grammar in Communicating Autonomous Robotic Agents. In Horn, W. (ed.) Proceedings of ECAI 2000, 
IOS Publishing, Amsterdam, pp 764-769.

@@

34.\citet{Steels:00f} Origine et \'evolution du langage : exp\'eriences robotiques. Revue du Palais de la D\'ecouverte, 278:63-67 May 2000.

@@

63. \citet{Steels:05b} Coordinating Perceptually Grounded Categories through Language. A Case Study for Colour. Target article. Behavioral and Brain Sciences. 28 (4). 469-490.

@@

66. \citet{Steels:05e} Planning What To Say: Second Order Semantics for Fluid Construction Grammars. 
In: Bugarin Diz, A. and J. Santos Reyes (eds.) Proceedings of CAEPIA 2005. Springer-Verlag, Berlin.

@@

74. \citet{Loreto:07a} Social Dynamics: Emergence of Language. Nature Physics 3(11) pp. 758-760.

@@

\clearpage
81. \citet{Steels:07h} Is Symbolic Inheritance Similar to Genetic Inheritance?. Behavioral and Brain Sciences, 30(4):376-377.

82. \citet{Steels:07i} Embodiment and Self-Organization of Human Categories: A Case Study for Speech. In: Ziemke, T. and Zlatev, J. and R. Frank (eds.) Body, Language, and Mind, Cognitive Linguistics Research, Mouton De Gruyter. Berlin, pp. 411-430.

83. \citet{Wellens:08a} Flexible Word Meaning in Embodied Agents. Connection Science, 20(2-3): 173-191.

@@

85. \citet{Steels:08c} The Robot in the Mirror. Connection Science 20(2-3), 337-358.

@@

100. \citet{Steels:09h} Babel. In: Nolfi, S. and M. Mirolli (eds) (2009) Evolution of Communication and Language in Embodied Agents. Springer Verlag, Berlin. 307-313.

@@

113. \citet{Steels:11d} Design Patterns in Fluid Construction Grammar. John Benjamins Pub. Amsterdam

@@

124. \citet{Spranger:12g} A perceptual system for language game experiments. In: In: Steels, L. and M. Hild (eds.) (2012) Language Grounding in Robots. Springer-Verlag, New York. pp. 95-116.

125. \citet{Spranger:12h} Open-ended Procedural Semantics. In: Steels, L. and M. Hild (eds.) (2012) Language Grounding in Robots. Springer-Verlag, New York. pp. 159-178.

@@

127. \citet{Steels:12j} Emergent action language on real robots. In: Steels, L. and M. Hild (eds.) (2012) Language Grounding in Robots. Springer-Verlag, New York. pp. 261-282.

@@

130. \citet{Steels:12m} Emergent mirror systems for body language. In: Steels, L. (ed) Experiments in Cultural Language Evolution. John Benjamins Pub., Amsterdam. pp. 87-109.

@@

135. \citet{Beuls:13a} Agent-Based Models of Strategies for the Emergence and Evolution of Grammatical Agreement. PLOS ONE, 8(3), e58960. \url{http://dx.plos.org/10.1371/journal.pone.0058960}

>>>>

\addchap{List of relevant papers by Luc Steels}
\label{c:references}

@@

1. \citet{Steels:1995} The `artificial life' route to `artificial intelligence'. Building Situated Embodied Agents. Lawrence Erlbaum Ass, New Haven, 1994.

@@

9. \citet{Steels:1998} The Origins of Syntax in Visually Grounded Robotic Agents. In Pollack, M. (ed.) Proceedings of the 15th International Joint Conference on Artificial Intelligence, Morgan Kaufmann Publishers, San Francisco, CA. pp. 1632-1641.

@@

\enlargethispage{1em}
13. \citet{Steels:97g} Ancrage de Jeux de Langage Adaptatifs dans des Agents Robotiques. In A. Cherif. J. and Signorini (ed.) Intelligence Artificielle et Complexite, Paris, 1997. pp. 27-39.

@@

17. \citet{Steels:1998} The origins of syntax in visually grounded robotic agents. Artificial Intelligence, 103:1-24.

@@

21. \citet{McIntyre:1999} Net-mobile embodied agents. Proceedings of Sony Research Forum, Tokyo, 1999.

@@

28. \citet{Steels:1999} Collective learning and semiotic dynamics. In Floreano, D. and 
Nicoud, J-D and Mondada, F., editor, Advances in Artificial Life (ECAL 99) LNCS 1674, Springer Verlag, Berlin. 
pp. 679-688.

@@

\newpage
31. \citet{Steels:00c} The Emergence of Grammar in Communicating Autonomous Robotic Agents. In Horn, W. (ed.) Proceedings of ECAI 2000, 
IOS Publishing, Amsterdam, pp 764-769.

@@

34.\citet{Steels:2000kaplan} Origine et \'evolution du langage : exp\'eriences robotiques. Revue du Palais de la D\'ecouverte, 278:63-67 May 2000.

@@

63. \citet{Steels:2005} Coordinating Perceptually Grounded Categories through Language. A Case Study for Colour. Target article. Behavioral and Brain Sciences. 28 (4). 469-490.

@@

\enlargethispage{1em}
66. \citet{Steels:05e} Planning What To Say: Second Order Semantics for Fluid Construction Grammars. 
In: Bugarin Diz, A. and J. Santos Reyes (eds.) Proceedings of CAEPIA 2005. Springer-Verlag, Berlin.

@@

74. \citet{Loreto:2007} Social Dynamics: Emergence of Language. Nature Physics 3(11) pp. 758-760.

@@

81. \citet{Steels:07h} Is Symbolic Inheritance Similar to Genetic Inheritance?. Behavioral and Brain Sciences, 30(4):376-377.

\enlargethispage{2em}
82. \citet{Steels:07i} Embodiment and Self-Organization of Human Categories: A Case Study for Speech. In: Ziemke, T. and Zlatev, J. and R. Frank (eds.) Body, Language, and Mind, Cognitive Linguistics Research, Mouton De Gruyter. Berlin, pp. 411-430.

83. \citet{Wellens:2008} Flexible Word Meaning in Embodied Agents. Connection Science, 20(2-3): 173-191.

@@

85. \citet{Steels:2008} The Robot in the Mirror. Connection Science 20(2-3), 337-358.

@@

% 9h is identical to 10e

@@

\newpage
113. \citet{Steels:11d} Design Patterns in Fluid Construction Grammar. John Benjamins Pub. Amsterdam

@@

124. \citet{Spranger:2012vision} A perceptual system for language game experiments. In: In: Steels, L. and M. Hild (eds.) (2012) Language Grounding in Robots. Springer-Verlag, New York. pp. 95-116.

125. \citet{Spranger:2012} Open-ended Procedural Semantics. In: Steels, L. and M. Hild (eds.) (2012) Language Grounding in Robots. Springer-Verlag, New York. pp. 159-178.

@@

127. \citet{Steels:2012b} Emergent action language on real robots. In: Steels, L. and M. Hild (eds.) (2012) Language Grounding in Robots. Springer-Verlag, New York. pp. 261-282.

@@

\newpage
130. \citet{Steels:12m} Emergent mirror systems for body language. In: Steels, L. (ed) Experiments in Cultural Language Evolution. John Benjamins Pub., Amsterdam. pp. 87-109.

@@

135. \citet{Beuls:2013} Agent-Based Models of Strategies for the Emergence and Evolution of Grammatical Agreement. PLOS ONE, 8(3), e58960. \url{http://dx.plos.org/10.1371/journal.pone.0058960}
